Case study validations of automatic bullet matching

### 2018-10-26 index.Rmd
phrasing
fix compile issues

labels <- h44nest %>% filter(samesource, sam_rf < 0.5)
labels$comparison <- with(labels, paste(bullet1, "vs", paste(barrel2, bullet2, sep = "-")))
labels$measure <- "RF score"
labels$study <- "Hamby 44"
labels$y <- 5 + 4:1
nestslong %>%
  ggplot(aes(
    x = value, fill = samesource,
    colour = samesource
  )) +
  geom_density(alpha = 0.6) +
  geom_rug(alpha = 0.5, show.legend = F) +
  facet_grid(measure ~ study) +
  theme_bw() +
  scale_fill_manual("Same source", values = c("darkgrey", "darkorange")) +
  scale_colour_manual("Same source", values = c("darkgrey", "darkorange")) +
  scale_x_continuous("SAM scores", limits = c(0, 1)) +
  theme(legend.position = "bottom") +
  geom_hline(yintercept = 0, show.legend = F) +
  geom_text(aes(x = sam_rf, y = y, label = comparison),
            show.legend = F,
            hjust = 0,
            data = labels
  ) +
  geom_segment(aes(x = sam_rf, xend = sam_rf, y = 0, yend = y),
               show.legend = F,
               size = 0.1,
               colour = "grey20",
               data = labels
  )
```

@@

```{r, cache=TRUE}
hounest <- hounest %>% mutate(
  data = data %>% purrr::map(
    .f = function(d) {
      d$sameland = bullet_to_land_predict(
        land1 = d$land1, land2 = d$land2, 
        d$rfscore, difference=0.05)
      d
    })
)
houlands <- hounest %>% unnest(data)

h44nest <- h44nest %>% mutate(
  data = data %>% purrr::map(
    .f = function(d) {
      d$sameland = bullet_to_land_predict(
        land1 = d$land1, land2 = d$land2, 
        d$rfscore, difference=0.1)
      d
    })
)
h44lands <- h44nest %>% unnest(data)

pdnest <- pdnest %>% mutate(
  data = data %>% purrr::map(
    .f = function(d) {
      d$sameland = bullet_to_land_predict(
        land1 = d$land1, land2 = d$land2, 
        d$rfscore, difference=0.1)
      d
    })
)
pdlands <- pdnest %>% unnest(data)

all_lands <- rbind(
  h44lands %>%
    select(rfscore, ccf, sameland, samesource) %>%
    mutate(study = "Hamby 44", group1 = ""),
  pdlands %>%
    select(rfscore, ccf, sameland, samesource) %>%
    mutate(study = "Phoenix PD", group1 = ""),
  houlands %>%
    select(rfscore, ccf, sameland, samesource, group1) %>%
    mutate(study = "Houston FSI")
)
all_lands$study <- factor(all_lands$study, levels = c("Hamby 44", "Phoenix PD", "Houston FSI"))
all_lands <- all_lands %>%
  mutate(
    sameland = replace(sameland, !samesource, FALSE)
  )
```

@@

```{r aucs, eval = FALSE}
aucs <- all_lands %>% group_by(study, group1) %>% nest() %>%
  mutate(
    auc_rf = data %>% purrr::map_dbl(function(d) {
      auc(d$sameland, d$rfscore)
    }),
    auc_ccf = data %>% purrr::map_dbl(function(d) {
      auc(d$sameland, d$ccf)
    })

  )
#
library(xtable)
names(aucs)[2] <- "set"
names(aucs)[4] <- "RF score"
names(aucs)[5] <- "Cross-correlation"
print(xtable(aucs %>% select(-data), digits = 3, caption = "Area under the curve (AUC) for each (set) of the studies. At the land level there is no difference in prediction power between RF score and cross-correlation."), include.rownames=FALSE, label="tab:aucs")
```

\begin{table}[ht]
\centering
\begin{tabular}{llrr}
  \hline
\bf Study & \bf Set & \bf AUC(RF score) & \bf AUC(Cross-correlation) \\ 
  \hline
\bf Hamby 44 &  & 0.956 & 0.926 \\ 
\bf   Phoenix PD &  & 0.976 & 0.986 \\ 
\bf   Houston FSI & G1 & 0.991 & 0.988 \\ 
\bf   Houston FSI & G2 & 0.898 & 0.902 \\ 
\bf   Houston FSI & G3 & 0.986 & 0.987 \\ 
   \hline
\end{tabular}
\caption{Area under the curve (AUC) for each (set) of the studies. As seen in the ROC curves, the algorithm performs the least well on set 2 of the Houston FSI study. Overall the performance on all sets is very good to excellent.
At the land level there is no difference in the predictive power between RF score and cross-correlation. \label{tab:aucs}} 
\end{table}

@@

<!-- \hh{XXX error rates on land to land matches? ROC curves in \autoref{fig:rocs} and corresponding AUC results for all studies in \autoref{tab:aucs} XXX} -->

@@

@@

What is quite surprising is that what we anticipated to be the easiest case of Hamby 44 turned out to be the hardest to solve. This is likely due to several lands in Hamby set 44 that have major deficiencies,  such as 'tank rash' (a collision of the bullet after exiting the barrel with a surface causing markings on top of the striations from the barrel) or extreme pitting (holes caused by direct contact with burning gun powder).
\autoref{fig:tankrash} gives an overview of all lands of affected bullets in the Hamby-44 set. These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent match scores.
```{r tankrash, fig.cap="Overview of bullet lands with prominent deficiency such as tank rash or extreme pitting in the Hamby-44 study.", fig.width=4, fig.height = 2, out.width='47.5%', fig.subcap=c("Br1-B1-L6 (tank rask)", "Br2-B2-L5 (tank rash)", "Br3-B1-L5 (tank rash)", "Br8-B1-L6 (tank rash)", "Br8-B2-L2 (pitting)", "Br8-B2-L6 (tank rash)", "Unk-E-L6 (tank rash)", "Unk-I-L6 (tank rash and break-off)"), fig.align="middle", fig.ncol=2}
knitr::include_graphics(
  c(
    "./images/B1-B1-L6.png", "./images/B2-B2-L5.png",
    "./images/B3-B1-L5.png", "./images/B8-B1-L6.png",
    "./images/B8-B2-L2.png", "./images/B8-B2-L6.png",
    "./images/NA-BE-L6.png", "./images/NA-BI-L6.png"
  )
)
```

@@

As seen in the earlier discussion, similarity scores are affected by deficiencies on lands. It has been shown \citet{lpr} that parts of lands can be used for successful identifications. The algorithm proposed by \citet{aoas} is not capable of  automatically detecting parts of lands with well expressed striae. \svp{It may be useful to couple the \citet{aoas} algorithm with an algorithm which assesses the quality of the input data and determines which portions of the data to use for comparison. This would emulate the process used by examiners, who first assess the quality of the evidence and whether there is enough information present to attempt a match.} XXX citation for neural network algorithm \svp{Do you want to release this idea into the wild at this point? }

>>>>

labels <- h44nest %>% filter(samesource, sam_rf < 0.5)
labels$comparison <- with(labels, paste(bullet1, "vs", paste(barrel2, bullet2, sep = "-")))
labels$measure <- "RF score"
labels$study <- "Hamby 44"
labels$y <- 5 + 4:1
nestslong %>%
  ggplot(aes(
    x = value, fill = samesource,
    colour = samesource
  )) +
  geom_density(alpha = 0.6) +
  geom_rug(alpha = 0.5, show.legend = F) +
  facet_grid(measure ~ study) +
  theme_bw() +
  scale_fill_manual("Same source", values = c("darkgrey", "darkorange")) +
  scale_colour_manual("Same source", values = c("darkgrey", "darkorange")) +
  scale_x_continuous("SAM scores", limits=c(0,1)) +
  theme(legend.position = "bottom") +
  geom_hline(yintercept = 0, show.legend = F) +
  geom_text(aes(x = sam_rf, y = y, label = comparison),
            show.legend = F,
            hjust = 0,
            data = labels
  ) +
  geom_segment(aes(x = sam_rf, xend = sam_rf, y = 0, yend = y),
               show.legend = F,
               size = 0.1,
               colour = "grey20",
               data = labels
  )
```

@@

```{r, cache=FALSE}
hounest <- hounest %>% mutate(
  data = data %>% purrr::map(
    .f = function(d) {
      d$sameland = bullet_to_land_predict(
        land1 = d$land1, land2 = d$land2, 
        d$rfscore, difference=0.05)
      d
    })
)
houlands <- hounest %>% unnest(data)

set.seed(20140501)
h44nest$samesource <- h44nest$sam_rf > 0.3
h44nest <- h44nest %>% 
  mutate(
    data = purrr::map2(.x = data, .y = samesource, .f = function(d, y) {
      d$samesource <- y
      d
  }
  ))

h44nest <- h44nest %>% mutate(
  data = data %>% purrr::map(
    .f = function(d) {
      if (any(d$samesource)) {
      d$sameland = bullet_to_land_predict(
        land1 = d$land1, land2 = d$land2, 
        d$rfscore, difference=0.01)
      } else d$sameland <- FALSE
      d <- d %>% select(-samesource)
      d
    })
)
h44lands <- h44nest %>% unnest(data)

pdnest <- pdnest %>% mutate(
  data = data %>% purrr::map(
    .f = function(d) {
      d$sameland = bullet_to_land_predict(
        land1 = d$land1, land2 = d$land2, 
        d$rfscore, difference=0)
      d
    })
)
pdlands <- pdnest %>% unnest(data)

all_lands <- rbind(
  h44lands %>%
    select(rfscore, ccf, cms, sameland, samesource) %>%
    mutate(study = "Hamby 44", group1 = ""),
  pdlands %>%
    select(rfscore, ccf, cms, sameland, samesource) %>%
    mutate(study = "Phoenix PD", group1 = ""),
  houlands %>%
    select(rfscore, ccf, cms, sameland, samesource, group1) %>%
    mutate(study = "Houston FSI")
)
all_lands$study <- factor(all_lands$study, levels = c("Hamby 44", "Phoenix PD", "Houston FSI"))
all_lands <- all_lands %>%
  mutate(
    sameland = replace(sameland, !samesource, FALSE)
  )
```

```{r check44, eval=FALSE}
h44lands <- h44lands %>%
  mutate(
    sameland = replace(sameland, !samesource, FALSE)
  )
h44lands %>% 
  ggplot(aes(x = land1, y = land2, fill = ccf)) +
  geom_tile() +
  geom_tile(aes(colour = sameland), size = 1, 
            data = h44lands %>% filter(sameland)) +
  scale_fill_gradient2("RF score: ", midpoint=0.5, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
  scale_colour_manual(values="darkorange") +
  facet_grid(bullet1~barrel2+bullet2)

```

```{r checkpd, eval=FALSE}
pdlands <- pdlands %>%
  mutate(
    sameland = replace(sameland, !samesource, FALSE)
  )
pdlands %>% 
  ggplot(aes(x = land1, y = land2, fill = ccf)) +
  geom_tile() +
  geom_tile(aes(colour = sameland), size = 1, 
            data = pdlands %>% filter(sameland)) +
  scale_fill_gradient2("RF score: ", midpoint=0.5, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
  scale_colour_manual(values="darkorange") +
  facet_grid(bullet1~barrel2+bullet2)

```

```{r checkhou, eval=FALSE}
houlands <- houlands %>%
  mutate(
    sameland = replace(sameland, !samesource, FALSE)
  )
hou1 <- houlands %>% filter(group1==group2, group1=="G1")

hou1 %>%
  ggplot(aes(x = land1, y = land2, fill = ccf)) +
  geom_tile() +
  geom_tile(aes(colour = sameland), size = 1, 
            data = hou1 %>% filter(sameland)) +
  scale_fill_gradient2("RF score: ", midpoint=0.5, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
  scale_colour_manual(values="darkorange") +
  facet_grid(bullet1~barrel2+bullet2)

hou2 <- houlands %>% filter(group1==group2, group1=="G2")

hou2 %>%
  ggplot(aes(x = land1, y = land2, fill = ccf)) +
  geom_tile() +
  geom_tile(aes(colour = sameland), size = 1, 
            data = hou2 %>% filter(sameland)) +
  scale_fill_gradient2("RF score: ", midpoint=0.5, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
  scale_colour_manual(values="darkorange") +
  facet_grid(bullet1~barrel2+bullet2)

hou3 <- houlands %>% filter(group1==group2, group1=="G3")

hou3 %>%
  ggplot(aes(x = land1, y = land2, fill = ccf)) +
  geom_tile() +
  geom_tile(aes(colour = sameland), size = 1, 
            data = hou3 %>% filter(sameland)) +
  scale_fill_gradient2("RF score: ", midpoint=0.5, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
  scale_colour_manual(values="darkorange") +
  facet_grid(bullet1~barrel2+bullet2)  
```

@@

```{r aucs, eval = FALSE, results = 'asis'}

aucs <- all_lands %>% group_by(study, group1) %>% nest() %>%
  mutate(
    auc_rf = data %>% purrr::map_dbl(function(d) {
      auc(d$sameland, d$rfscore)
    }),
    auc_ccf = data %>% purrr::map_dbl(function(d) {
      auc(d$sameland, d$ccf)
    }),
    auc_cms = data %>% purrr::map_dbl(function(d) {
      auc(d$sameland, d$cms)
    }),
    auc_rf_ci_low = data %>% purrr::map_dbl(function(d) {
      ci.auc(d$sameland, d$rfscore)[1]
    }),
    auc_rf_ci_upper = data %>% purrr::map_dbl(function(d) {
      ci.auc(d$sameland, d$rfscore)[3]
    }),
    auc_ccf_ci_lower = data %>% purrr::map_dbl(function(d) {
      ci.auc(d$sameland, d$ccf)[1]
    }),
    auc_ccf_ci_upper = data %>% purrr::map_dbl(function(d) {
      ci.auc(d$sameland, d$ccf)[3]
    }),
    auc_cms_ci_lower = data %>% purrr::map_dbl(function(d) {
      ci.auc(d$sameland, d$cms)[1]
    }),
    auc_cms_ci_upper = data %>% purrr::map_dbl(function(d) {
      ci.auc(d$sameland, d$cms)[3]
    })

  )
#
library(xtable)
names(aucs)[2] <- "set"
names(aucs)[4] <- "RF score"
names(aucs)[5] <- "Cross-correlation"
names(aucs)[6] <- "CMS"

aucs <- as.data.frame(aucs)
aucs[,4] <- sprintf("%.3f (%.3f, %.3f)", aucs[,4], aucs[,7], aucs[,8])
aucs[,5] <- sprintf("%.3f (%.3f, %.3f)", aucs[,5], aucs[,9], aucs[,10])
aucs[,6] <- sprintf("%.3f (%.3f, %.3f)", aucs[,6], aucs[,11], aucs[,12])
print(xtable(aucs[,1:6] %>% select(-data), digits = 3, caption = "Area under the curve (AUC) for each (set) of the studies. At the land level there is no difference in prediction power between RF score and cross-correlation."), include.rownames=FALSE, label="tab:aucs")
```

\begin{table}[ht]
\centering
\begin{tabular}{llrrr}
  \hline
 && \multicolumn{3}{l}{\bf AUC (95\% Conf. Int)}\\
\bf Study & \bf Set & \bf RF score & \bf Cross-correlation & \bf CMS \\
  \hline
\bf Hamby 44 &  & 0.937 (0.917, 0.957) & 0.916 (0.891, 0.941) & 0.820 (0.787, 0.854) \\
\bf  Phoenix PD &  & 0.976 (0.957, 0.994) & 0.986 (0.975, 0.996) & 0.879 (0.838, 0.919) \\
\bf  Houston FSI & Set 1 & 0.991 (0.983, 0.999) & 0.988 (0.979, 0.997) & 0.785 (0.727, 0.842) \\
\bf  Houston FSI & Set 2 & 0.898 (0.853, 0.943) & 0.902 (0.861, 0.943) & 0.749 (0.684, 0.814) \\
\bf  Houston FSI & Set 3 & 0.986 (0.973, 0.999) & 0.987 (0.978, 0.997) & 0.911 (0.880, 0.941) \\
   \hline
\end{tabular}
\caption{Area under the curve (AUC) for each (set) of the studies for the random forest score, cross-correlation and CMS. Overall the perform on all sets is very good to excellent for cross-correlation and the random forest score.  As seen in the ROC curves, the algorithm performs the least well on set 2 of the Houston FSI study.  At the land level there is no significant difference in prediction power between RF score and cross-correlation.} \label{tab:aucs}
\end{table}

@@

@@

```{r why, fig.cap="Density plots for all measures "}
all_lands %>% gather(measure, score, ccf, cms, rfscore) %>%
ggplot(aes(x = score, fill = interaction(measure, sameland))) + 
  geom_density(alpha = 0.5) +
  scale_fill_brewer(palette = "Paired") +
  facet_grid(measure~study+group1, scales="free_x") 
```

@@

What is quite surprising is that \svp{while} we anticipated \svp{Hamby 44 would be the easiest set to work with, but it} turned out to be the hardest to solve. This is likely due to several lands in Hamby set 44 that have major deficiencies,  such as 'tank rash' (a collision of the bullet after exiting the barrel with a surface causing markings on top of the striations from the barrel) or extreme pitting (holes caused by direct contact with burning gun powder).
\autoref{fig:tankrash} gives an overview of all lands of affected bullets in the Hamby-44 set. These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent match scores.
```{r tankrash, fig.cap="Overview of bullet lands with prominent deficiency such as tank rash or extreme pitting in the Hamby-44 study.", fig.width=4, fig.height = 2, out.width='47.5%', fig.subcap=c("Br1-B1-L6 (tank rask)", "Br2-B2-L5 (tank rash)", "Br3-B1-L5 (tank rash)", "Br8-B1-L6 (tank rash)", "Br8-B2-L2 (pitting)", "Br8-B2-L6 (tank rash)", "Unk-E-L6 (tank rash)", "Unk-I-L6 (tank rash and break-off)"), fig.align="middle", fig.ncol=2}
knitr::include_graphics(
  c(
    "./images/B1-B1-L6.png", "./images/B2-B2-L5.png",
    "./images/B3-B1-L5.png", "./images/B8-B1-L6.png",
    "./images/B8-B2-L2.png", "./images/B8-B2-L6.png",
    "./images/NA-BE-L6.png", "./images/NA-BI-L6.png"
  )
)
```

@@

As seen in the earlier discussion, similarity scores are affected by deficiencies on lands. It has been shown \citet{lpr} that parts of lands can be used for successful identifications. The algorithm proposed by \citet{aoas} is not capable of  automatically detecting parts of lands with well expressed striae. \svp{It may be useful to couple the \citet{aoas} algorithm with an algorithm which assesses the quality of the input data and determines which portions of the data to use for comparison. This would emulate the process used by examiners, who first assess the quality of the evidence and whether there is enough information present to attempt a match.}
